---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into the following options - under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus - and calculates how many loans are in each group. How would you describe the distribution of loans in a sentence?
Hint: You will need to use `case_when` to make this work.
**A1** The majority of loans, 24000, are in the 0-50k price range, with over 15000 loans in the 10-50k range.

```{r}
# Write the code to produce your answer here
west_virginia_loan_amounts <- west_virginia_ppp %>%
mutate(
  amount_category = case_when(
    amount < 10000 ~ 'under_10k',
    amount >= 10000 & amount < 50000 ~ '10k_50k',
    amount >= 50000 & amount < 100000 ~ '50k_100k',
    amount >= 100000 & amount < 250000 ~ '100k_500k',
    amount >= 250000 & amount < 500000 ~ '500k_1m',
    amount >= 1000000 ~ '1m_plus')
  )
west_virginia_loan_amounts %>%
  group_by(amount_category) %>%
  summarise(
    count_loans = n()
  )
```

**Q2** Some of the loans appear not to have been paid out - there is a column representing the amount of money that hasn't been disbursed to the applicant. Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Davis Trust Company have the most with 12 loans, Miners & Merchants Bank have 1

```{r}
# Write the code to produce your answer here
west_virginia_undisbursed_amount <- west_virginia_ppp %>% filter(undisbursed_amount > 0)
  
west_virginia_undisbursed_amount %>%
    group_by(servicing_lender_name) %>%
    summarise(total= n()
    )
```

**Q3** What are some potential reasons why a loan might not be paid out? What questions does that circumstance raise for you?
**A3** Trouble with paperwork, something comes up after being approved for the loan that makes that company no longer eligible. Additionally, it a bit suspicious of the lender. In our data, only 13 laons are unpaid out, and 12 of those are from the same lender. Makes me suspicious of Davis Trust Company. 


**Q4** Which month in 2021 had the most loans approved?
**A4** Looks like in 2021 it was March. Not sure how to order by the date, so will need help on that one. 

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>% 
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(date_approved) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

**Q5** How many loans went to West Virginia applicants that listed the maximum number of jobs retained (500)? Which one had the earliest approved date?
**A5** 10, J. AUSTIN HEATING & COOLING, INC. on April, 8 2020

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>% 
  filter(jobs_retained == 500 )
    
```
