---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.


## Things I learned on Jan. 27

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```

## Things I learnd on Feb. 1

Lubridate is an R package that makes it easier to work with dates and times (part of the tidyverse)

Mutate code means add a new column, add or replace values in a column that exists or a new one we are making. 
Done by the code "mutate(name_of_column) = value)". Can name code anything you want but make it meaningful

You can then add an arrange column to arrange the new column in descending (or ascending order)

function "filter()" is a tidyverse verb that uses criteria you supply to return a subset of the original data. Can help you search for something within a column. "filter(column_name == 'Thing im looking for'), `filter(COLUMN_NAME OPERATOR CRITERIA)

function "if_else()" statements used similarly to Javascript

function "case_when()" helps group things together (put things in different buckets, as prof said, see task 9 prelab)

Use glimpse to check if a column is actully a date column, if not, use lubridate to make them dates (task 14 pre lab 02)




