---
title: "lab_03"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
```

## Load data
```{r}
# Write code to load the WV slice of the ppp loan data here and call it wv_ppp_loans.
# The data is called lab_03.rds, and is in the same folder as lab_03.Rmd (this data notebook)

wv_ppp_loans <- read_rds("lab_03.rds")

```

## Answer questions

Q1: Which five cities located in Kanawha County, West Virginia, got the largest _number_ of loans? How many did each get? To answer this question, you'll have to clean the very messy "city" field.

Note: To answer this question, you'll need to see every variation of the "city" column in Kanawha County, and then fully clean the city field for those cities. Hint: once cleaned, the dataframe that counts the number of loans by city should have 38 distinct cities for Kanawha County.

A1: Charleston (2802), South Charleston (478), Saint Albans (296), Dunbar (162), and Nitro (140)
```{r}

# Write your code in this codeblock.
cleaned_wv_ppp_loans <- wv_ppp_loans %>% filter(project_county_name == "KANAWHA") %>%
  clean_names() %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(city = str_to_title(city)) %>% 
  mutate(city = case_when(
    city == "Balle" ~ "Belle",
    city == "Charelston" ~ "Charleston",
    city == "Crosslanes" ~ "Cross Lanes",
    city == "Elkveiw" ~ "Elkview",
    city == "S Charleston" ~ "South Charleston",
    city == "So Charleston" ~ "South Charleston",
    city == "St Albans" ~ "Saint Albans",
    city == "St.albans" ~ "Saint Albans",
    city == "St. Albans" ~ "Saint Albans",
    TRUE ~ city
    ))
 
  
cleaned_wv_ppp_loans %>%
  group_by(city) %>%
  summarise(total = n()) %>%
  arrange(desc(total))


```

Q2: Find the 5-digit ZIP code in West Virginia that had the highest total amount of loans (by dollar value). What is the name of the business that got the largest loan (by dollar value) in that zip code? Write up a short paragraph that also answers these questions:

* Where in West Virginia is this ZIP code located?
* How much was the business' loan for?
* How much of that loan was forgiven?
* Which number is bigger, the loan amount or the forgiveness amount? Can you explain the discrepancy?

A2: 25301 is the highest total amount for a zip code. I'm not sure what business it is that has the largest loan (couldn't get that part of the code) but I believe it is NATIONAL TRAVEL SERVICE INC. The loan was for 788,877 dollars. But they have an additional loan under a slightly different zip code that is also 749,150 dollars. I'm not sure how much of the loan has been forgiven. The zip code is in Charleston is Kanawha county. I don't see a forgiveness amount listed. 

```{r}
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount
# Second, use the top ZIP code to create a separate dataframe that shows the business with the biggest loan in that ZIP code.
zip_wv_ppp <- wv_ppp_loans %>%
  clean_names() %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) 
  
zip_wv_ppp %>% 
 group_by(zip) %>%
 summarise(amount = sum(amount)) %>%
  arrange(desc(amount))
  
new_zip <- wv_ppp_loans %>% filter(zip == "25301") %>%
clean_names() %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
summarise(total = n())
  
```

Q3: Five business entities owned by the spouse of a member of West Virginia's U.S. House delegation got PPP loans. Write up a paragraph that answers the following questions.   

  * What are the businesses' names?
  * What do the businesses do?
  * How much money did the five businesses receive in total?
  * What percentage of the total loan amount have been forgiven?
  * How did the U.S. House member vote on the legislation initially author zing the PPP loan program.

Feel free to add any interesting "color" (little interesting details) you find in your research.  

Some hints to help you answer this question:

* West Virginia's three U.S. House members are listed [here](https://www.govtrack.us/congress/members/WV#representatives)
* To find businesses they or their families own, you can search for financial disclosure forms for U.S. House members [here](https://disclosures-clerk.house.gov/PublicDisclosure/FinancialDisclosure#Search).
  * To speed things up, here are direct links to the three U.S. House members most recent financial disclosures, which I found by searching the site linked above.
  * [Rep. 1](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042042.pdf)
  * [Rep. 2](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10040330.pdf)
  * [Rep. 3](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042734.pdf)
* When you examine the disclosure forms, there may be lots of stock sales and retirement accounts mixed in.  Ignore those. In the table in the disclosure forms, look for "Partnership" or "C-Corp" in the column named "Income Type"
* This is a link to details on passage of the [CARES ACT](https://www.congress.gov/bill/116th-congress/house-bill/748). You can find a link to how the House voted by clicking on a link from [this page](https://www.congress.gov/bill/116th-congress/house-bill/748/actions).

A3: Business names include DUTCH MILLER OF CHARLESTON, INC, DUTCH MILLER OF CARLOTTE, INC, DUTCH MILLER SUBARU, INC, DUTCH MILLER CHEVROLET, INC., and DM MOTOR, INC. They are all car dealerships and sell new and used cars. They recieved a total of 3,105,300 dollars. All of the loans have been paid in full. She voted for the program, once saying "I voted for the Paycheck Protection Program and Healthcare Enhancement Act to deliver much-needed small business relief, keep Americans employed, send resources to frontline providers and hospitals and expand COVID-19 testing." I couldn't figure out how to filter out the data by exact company name. I tried a filter and thought it might be a case(when) but I couldn't figure it out. I ddi research and went through thr wider range of data to try to answer as many of the questions as I could regardless.

```{r}

# Hint
# First, create a dataframe that only has loans to those five companies
# Second, use that newly created dataframe to calculate total loan amount and percentage forgiven for those companies

five_businesses <- wv_ppp_loans %>% 
filter(df, date %>%  c("DUTCH MILLER OF CHARLESTON, INC", "DUTCH MILLER OF CHARLOTTE, INC"))
```
