---
title: "Data Analysis Project"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Will be looking at statistics from this paragrapph "For most of the decade before 2015, Baltimore’s annual homicide arrest rate hovered at about 40 percent. Since 2015, the arrest rate hasn’t topped 30 percent in any year. And while most cities saw their arrest rates drop gradually, Baltimore’s decline was sudden — plummeting 15 percentage points in 2015, after Gray’s death, the largest single-year drop for any city already solving less than half its homicides." 



```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```


## read in the file from github 

```{r}
library(readr)

urlfile="https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicide_data <- read_csv(url(urlfile))

```

##filter the file to only inlcude data from Baltimore

##then filter that data to only look at crimes reported before 2015

```{r}
baltimore_homicide_data <- homicide_data %>% 
  filter(city == "Baltimore")

early_baltimore_homicide_data <- baltimore_homicide_data %>% 
  filter(reported_date < 20150103)
```

"
## find arress rate from "most of tghe decade before 2015" to be 40%

```{r}
successful_arrests <- early_baltimore_homicide_data %>% 
  filter(str_detect(disposition ,'by arrest'))





```


## We find from the 1,825 cases in  Baltimore before 2015, that 750 were closed by arrest. This means that before 2015, roughly 41% of cases ended in an arrest which checks out with the "For most of the decade before 2015, Baltimore’s annual homicide arrest rate hovered at about 40 percent" sentence.


##next I will code to find the findings of the next sentence



