---
title: "state_profile"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PPP State Profile

Your assignment is to explore a dataset of PPP loan applications for a particular state and to answer a series of questions that will help you summarize that state's loan applications. You will need to write R code to answer those questions. You also will bring in other data, including Census information, and you will be making graphics, including maps, to illustrate your findings.

The deliverable will be this R Markdown notebook and a data folder that you will receive with a state's loan application data. Place any other data you are using to answer the questions in the same data folder.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this assignment.

```{r}

library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library(tigris)
library(corrr)
library(sf)
library(tidycensus)
census_api_key("72dab2f9b69151aab34c4c94ddc8a797a04f3944")

colorado_ppp <- read_csv("data/colorado_ppp.csv")
naics_codes <- read_csv("data/naics_codes.csv")
```

## Initial Exploration

**Q1.**. Write R code that generates some basic descriptive statistics that allows you to describe the applications from your state. This code should produce results that describe the data in different ways: how many applications there are, along with ways to understand the typical and most frequent values for columns you find interesting or newsworthy. You should produce at least five (5) descriptive statistics and write up a summary of the data using them.

**A1.** There were 192,266 loan applications recorded in Colorado with loan amounts totaling $15,059,968,209 dollars. The top 4 counties in Colorado with the most applications are Denver, Arapahoe, Jefferson, and El Paso (in that order) which is no surprise, as these counties are also the four most populated counties in Colorado. The highest loan averages by county come from Denver, followed by Arapahoe, Adams, Boulder, and Pitkin. Business wise, the most loans in Colorado go to Full-Service Restaurants, followed by Offices of Real Estate Agents and Brokers, Beauty Salons, Beef Cattle Ranching and Farming, and Offices of Dentists. The largest amount of money in loans are distributed to Full-Service Restaurants, Offices of Physicians, Hotels (not casino hotels) and motels, Limited-Service Restaurants, and Offices of Dentists. The max amount for a loan application is 10 million dollars, which only 13 companies applied for. And in Colorado, 15.6% of applicants were White, 2% were Asian, 1.4% were Black or African American, 0.3% were American Indian or Alaksa Native, 0.06% were Native Hawaiian and 0.002% were Puerto Rican. 80.6% of all applicants failed to declare their race. Every single loan in the dataframe has been disbursed, and the loans are predominately from Urban areas, with only 13.8% of loans being identified as Rural. 


```{r}
colorado_ppp %>% 
  group_by(state) %>% 
  summarize(total_loans=n())

## There are 192,266 loans in Colorado

colorado_ppp %>% 
   summarise(total_amount = sum(amount))
## The loans add up to $15,059,968,209 dollars

loan_by_county <- colorado_ppp %>% 
  group_by(project_county_name) %>% 
  summarize(total_loans=n()) %>% 
  arrange(desc(total_loans))
## The top 4 counties with the most loans are Denver, Arapahoe, Jefferson, and El Paso (in that order). These counties are also the four most populated counties in Colorado (in the order of Denver County, El Paso County, Arapahoe County, and Jefferson County).

loan_by_naics <- colorado_ppp %>% 
  group_by(naics_code) %>% 
  summarize(total_amount=sum(amount), total_loans=n()) %>% 
  arrange(desc(total_loans))

top_five_naics <- naics_codes %>% 
  filter(naics_codes == 722511 | naics_codes == 531210 | naics_codes == 812112 | naics_codes == 112111 | naics_codes == 621210)

## The most loans in Colorado go to Full-Service Restaurants, followed by Offices of Real Estate Agents and Brokers, Beauty Salons, Beef Cattle Ranching and Farming, and Offices of Dentists


total_amount_naics <- loan_by_naics %>% 
  arrange(desc(total_amount))

top_five_naics_amount <-naics_codes %>% 
  filter((naics_codes == 722511 | naics_codes == 621111 | naics_codes == 721110 | naics_codes == 722513 | naics_codes == 621210))

## The most total loan amount in Colorado goes to Full-Service Restaurants, Offices of Physicians, Hotels (not casino hotels) and motels, Limited-Service Restaurants, and Offices of Dentists. 


colorado_ppp %>% 
  arrange(desc(amount))

## The most a loan can be in Colorado is $10,000,000. Only 13 businesses receive loans of this amount, they were all in Denver, Jefferson or El Paso counties. The smallest loan given out was $46 to CELEBRATE SAFETY INC. 
  
mean_loan_ppp <- colorado_ppp %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount)
  ) %>%
  arrange(desc(mean_loan_amount))

## mean loan amount. Shows the data set has a couple stray counties, including a county from California and Counties from Texas. The highest loan avergaes by county are Denver, Arapahoe, Adams, Boulder, and Pitkin. 

colorado_ppp %>% 
  count(race) %>% 
  mutate(
    percent = n / sum(n) * 100
  ) %>% 
  arrange(desc(n))

## Loans by race and ethnicity: 80.6% of all applicants failed to admit their race. Of those who did, 15.6% were white, 2% were asian, 1.4% were Black or African American, ).3% were American Indian or Alaksa Native, 0.06% were Native Hawaiian and 0.002% were Puerto Rican

colorado_ppp %>% 
  count(rural_urban_indicator) %>% 
  mutate(
    percent = n / sum(n) * 100
  )

## Loans rural v urban: 13.8 % rural and 86.1% urban

colorado_ppp %>% 
   summarize(
    total_amount_disbursed = sum(amount),
    total_amount_undisbursed = sum(undisbursed_amount)
  )

## disbursed vs undisbursed loans: 100% of loans are disbursed
```



## Geographic Analysis

**Q2.** Write R code that examines geographic patterns for PPP loans in your state, using Census population information to calculate a per-capita figure for the state and counties and zip codes. Then, make a county map using ggplot showing the per-capita data and a zip code map showing the difference from the statewide per-capita figure. Describe the most interesting or newsworthy findings based on your exploration.

**A2.** An interesting finding I came across was that the area code 80477, which is the city of Steamboat Springs, had over three times the ppp per capita amount than any other zip code. They registered $133,711.685 in amount per capita, with the second highest zip code only regsitering 38,785.062. I researched the city and found that it is home to a ski resort, which would explain the high per capita amount of loans. After all, the population in 80477 is only 76 people. So although only 76 people are registered as living here full-time, the town is overrun with travelers in the winter visiting the Steamboat ski resort and historic Old Town hot springs. 

```{r}
## clean up the zip codes so they are only 5 digits. Also converted to a numeric value. 
colorado_ppp <- colorado_ppp %>% 
  mutate(
    original_zip = zip) %>% 
  mutate(
    zip = str_sub(original_zip, start=1L, end=5L)) %>% 
  mutate(
    zip = as.numeric(zip))

## code to create per-capita figures for the state of Colorado
colorado_state_pop <- get_acs(geography = "state", 
              variables = c(population = "B01001_001"), 
              state = "CO",
              year = 2019,
              geometry = TRUE)
colorado_state_pop <- colorado_state_pop %>%
  rename(
    state = NAME,
    population = estimate
  ) %>%
  select(state, population)
co_whole_state_ppp <- colorado_ppp %>% 
  group_by(state) %>% 
  summarize(
    total_ppp_amount = sum(amount)
  ) 

co_whole_state_ppp$state <- str_replace(co_whole_state_ppp$state, "CO", "Colorado")

colorado_state_ppp_per_capita <- colorado_state_pop %>% 
  left_join(co_whole_state_ppp, by = c("state")) %>%
  mutate(
    per_capita_ppp_amount = total_ppp_amount / population)
  drop_na(population)

## code to create per capita figures by county
colorado_county_pop <- get_acs(geography = "county", 
              variables = c(population = "B01001_001"), 
              state = "CO",
              year = 2019,
              geometry = TRUE)
colorado_county_pop <- colorado_county_pop %>% 
  mutate(NAME = toupper(str_remove_all(NAME, " County, Colorado")))

colorado_county_pop <- colorado_county_pop %>% 
  rename(
    project_county_name = NAME,
    population = estimate
  ) %>% 
  select(project_county_name, population, geometry) 

colorado_county_pop
    
co_county_ppp <- colorado_ppp %>% 
  group_by(project_county_name) %>% 
  summarize(
    total_ppp_amount = sum(amount)
  ) 
colorado_county_ppp_per_capita <- colorado_county_pop %>% 
  left_join(co_county_ppp, by = c("project_county_name")) %>% 
  mutate(
    per_capita_ppp_amount = total_ppp_amount / population) %>% 
  drop_na(population)

#code for per-capita figures by zip code
colorado_zip_pop <- get_acs(geography = "zcta", 
              variables = c(population = "B01001_001"), 
              state = "CO",
              year = 2019,
              geometry = TRUE)
colorado_zip_pop <- colorado_zip_pop %>% 
  rename(
    zip = GEOID,
    population = estimate
  ) %>% 
  mutate(
    zip = as.numeric(zip)
    ) %>% 
  select(zip, population)
colorado_ppp_by_zip <- colorado_ppp %>% 
  group_by(zip) %>% 
  summarize(
    total_ppp_amount = sum(amount)
  )

colorado_ppp_by_zip_per_capita <- colorado_zip_pop %>% 
  left_join(colorado_ppp_by_zip, by = c("zip")) %>% 
  mutate(
    per_capita_ppp_amount = total_ppp_amount / population
  ) %>% 
  drop_na(population)

## county map gg plot using gg plot
cocounties = counties("CO", cb = TRUE)
ggplot(cocounties) + 
  geom_sf(data = colorado_county_ppp_per_capita, aes(fill =  per_capita_ppp_amount)) + 
  coord_sf(
    crs = st_crs(4269), 
    xlim = c(-109.0603, -102.0405), 
    ylim = c(36.99243, 41.00344), 
    expand = FALSE, 
    datum = NA
    ) +
  scale_fill_viridis_c(option = "Houston") +
  theme_minimal() +
  labs(title = "PPP loans per capita by County")
  
## zip codes map using gg plot
ggplot(cocounties) + 
  geom_sf(data = colorado_county_pop) + 
  coord_sf(
    crs = st_crs(4269), 
    xlim = c(-109.0603, -102.0405), 
    ylim = c(36.99243, 41.00344), 
    expand = FALSE, 
    datum = NA
    ) +
  geom_sf(data = colorado_zip_pop, aes(fill = population)) +
  coord_sf(
    crs = st_crs(4269), 
    xlim = c(-109.0603, -102.0405), 
    ylim = c(36.99243, 41.00344), 
    expand = FALSE, 
    datum = NA
    ) +
 scale_fill_viridis_c(option = "magma", trans = "log") +
  theme_minimal() +
  labs(title = "PPP loans per capita by five digit Zip Code")

```

## Lender Analysis

**Q3.** Write R code to examine which lenders had the most approved applications in your state (and include summary statistics such as total amount and average loan amount). Generate dataframes that show the number and total amount of all loans and undisbursed loans per lender. For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. Describe the most noticeable outlier lenders on that graphic below. If there are no undisbursed loans, make graphic showing the relationship between lenders and loan amounts showing how many loans each lender issued for each amount in the dataframe. Describe the most noticeable outlier lenders on that graphic.

**A3.** The most noticeable outlier lender in the group is the First Bank lender. They have given out 7,000 more loans than the next closest lender and are the only lender who have given over a billion dollars, at $1.4 billion, almost doubling JPMorgan Chase Bank, National Association who have given the second highest amount of money at 897 million. These two lenders as well as Wells Fargo Bank, National Association and U.S. Bank, National Association are the only lenders to give over 10,000 loans, so they stand out among the others. Another lender that stands out is Zions Bank, A Division Of who have only given out 6,020 loans but here loan amount is 686 million, making them the third highest lender by amount. An even bigger outlier is BOKF, National Association, who have a total loan amount of 670 million dollars, the fourth most, while only giving out 2,168 loans. Only 45 of 1,258 lenders have given over 1,000 loans. 

```{r}
lender_ppp <- colorado_ppp %>% 
  group_by(lender) %>% 
  summarize(
    total_loans = n(),
    amount_millions = sum(amount) / 1000000
  )%>% 
  arrange(desc(total_loans))

## First bank has given out way more total loans than anyone else, almost 7,000 more. They've also given also double the amount of the next lender, over 1 billion dollars

lender_ppp %>% 
  ggplot(aes(x=total_loans, y=amount_millions)) + geom_point() +
  labs(
    title="Relationship between lenders and loan amount",
    x = "Total Number of Loans",
    y = "Total Amount of Money Disbursed (millions)",
    caption = "source: data from SBA PPP loan database")

## All loans in Colorado are disbursed
```


## Industry Analysis

**Q4.** Write R code that examines industry patterns for PPP loans in your state, using the NAICS codes from the PPP data as a starting point. Generate statewide and county industry totals, then join that with 2018-19 data from the [Census County Business Patterns survey](https://www2.census.gov/programs-surveys/cbp/datasets/2019/cbp19co.zip) using 6-digit NAICS codes. The documentation explaining that data can be found here: https://www2.census.gov/programs-surveys/cbp/technical-documentation/records-layouts/2018_record_layouts/county-layout-2018.txt. To do this, you will need to add FIPS codes to your PPP dataset.

Does the distribution of PPP applications by the top 10 industries (by number of applications) roughly match the number of businesses reported in the Census data? Does it roughly match if you remove self-employed individuals and sole proprietorships from the PPP totals? Write up a summary of what you've found and whether there are potentially newsworthy patterns in the data.

Create a county-level map showing the differences between the PPP applications and the Census business data for one of the top 10 industry codes. You can do this either using ggplot or Datawrapper (if the latter, publish that map and include the URL below).

**A4.** The distribution of the top 10 industries does not match up with the number of businesses whatsoever. Even after removing self-employed individuals and sole proprietorship's it is still way off. While getting rid of the individualistic businesses does shrink the gaps, it is super marginal. This could mean that a lot of sole proprietorship business aren't reported to the census or the possibility that these businesses aren't actually self-employers. Three naics_codes have differences greater than 1000, even after eliminating self-employed individuals and sole proprietorship's. These discrepancies point to two things: businesses not being correctly reported or reported at all to the Census or the PPP loan data in Colorado has a lot of faults and errors. 

Other than that there are not many meaningful trends. The discrepancies were pretty large all around, so there isn't necessarily a particular business that seemed suspect, more that the loans themselves/collection of Census data were fraudulent or unreported. As for the map, Adams County and Arapahoe County clearly had the biggest discrepancies in data regarding Beauty Salons. What that means, I'm not entirely sure. 

```{r}
## read in provided zip
real_naics <- read_csv("data/cbp19co.txt") %>% 
  rename(
    fips_codes = fipstate,
    naics_code = naics
  ) %>% 
  filter(fips_codes == "08")

##begin to calculate statewide industry totals
statewide_industry <- colorado_ppp %>% 
  count(naics_code) %>% 
  mutate(
    naics_code = as.character(naics_code)
  ) %>% 
  arrange(desc(n)) %>% 
  head(10)

statewide_naics <- real_naics %>% 
  group_by(naics_code) %>% 
  summarize(
    total = sum(est))

statewide_industry <- statewide_industry %>% 
  left_join(statewide_naics, by = c ("naics_code")) 

statewide_industry <- statewide_industry %>% 
  mutate(
    difference = abs(n - total))

## The difference between the Census data and that of the loan dataset are very different, meaning one of them has to be incorrect or outdated. 

## totals, after removing all self-employed individuals and sole proprietorships

non_solo_statewide_industry <- colorado_ppp %>% 
  filter(business_type != "Self-Employed Individuals", business_type != "Sole Proprietorship") %>% 
  count(naics_code) %>% 
  mutate(
    naics_code = as.character(naics_code)) %>% 
  arrange(desc(n)) %>% 
  head(10)

non_solo_statewide_industry <- non_solo_statewide_industry %>% 
  left_join(statewide_naics, by = c ("naics_code")) 

non_solo_statewide_industry <- non_solo_statewide_industry %>% 
  mutate(
    difference = abs(n - total))

## Differences all shrunk, however, they are still large with three of them having differences > 1000
 
## County-level map for 812112
counties_naics <- real_naics %>% 
  filter(naics_code == "812112")
cocounties_naics <- cocounties %>% 
  rename(
    fipscty = COUNTYFP) %>% 
    left_join(counties_naics, by = c("fipscty"))
cocounties_naics <- cocounties_naics %>% 
  select(NAME, est) %>% 
  rename(
    county = NAME) %>% 
  mutate(
    county = str_to_upper(county))

industry_trends_by_county <- colorado_ppp %>% 
  mutate(
    naics_code = as.character(naics_code)) %>% 
  filter(naics_code == "812112") %>% 
  count(project_county_name) %>% 
  rename(county = project_county_name)

industry_trends_by_county <- industry_trends_by_county %>% 
  left_join(cocounties_naics, by = c("county"))

industry_trends_by_county <- industry_trends_by_county %>% 
  mutate(
    est = ifelse(is.na(est), 0, est)) %>% 
  mutate(
    data_difference = n - est) %>% 
relocate(geometry, .after = data_difference)
  

cocounties_naics <- cocounties_naics %>% 
  left_join(industry_trends_by_county)
ggplot(cocounties_naics) +
  geom_sf(data = cocounties_naics, aes(fill = data_difference)) +
   coord_sf(
    crs = st_crs(4269), 
    xlim = c(-109.0603, -102.0405), 
    ylim = c(36.99243, 41.00344), 
    expand = FALSE, 
    datum = NA
    ) +
 scale_fill_viridis_c(option = "magma") +
  theme_minimal() +
  labs(title = "Difference between PPP and Census data for Beauty Salons")
 
```

## Summary

**Q5.** What is the most promising story idea that you have found during this exploration? How would you approach reporting it out and what other information/data would you need to do that? What would be the minimum story you could get from this, and what would be the maximum story?

**A5.** A small story could be about how the PPP loans are dominated by the largely populated counties such as Denver and how you can clearly see the influence of skiing in areas like Boulder who have almost half the population of Adams and El Paso counties, but roughly the same amount of loans. If you wanted to use per capita information, you could delve back into the town of Steamboat Springs and look at how the skiing industry legitimately keeps a city of 76 people afloat. Big picture, you could look into the legitimacy of the loans as there were incredibly large discrepancies in the data from the PPP database and that of the Census. Specifically the Taxi Service industry who had a 99% discrepancy, looking into those kinds of businesses could allow you to dig up something seriously wrong in the loan system or something being abused by businesses. 
